\SW: II. New Gravitational Lens Candidates from the CFHTLS Discovered through Citizen Science

### 2015-05-09 doc/sw-cfhtls.tex
Rewording in discussion. Power seems to be more a property of a *classification* than it does an agent... Please merge and edit as you see fit, Anu! #203

\subsubsection{Blind lens search}
Efficiency of a visual search can vary in different sections of an
image. Our eyes tend to focus usually at the center of an image and lens
candidates close to the borders could go undetected. Therefore, it is
essential to test and understand if \sw is missing some of the known lenses
because they happen to be close to the borders of the image cutouts.

@@

We do not find any strong visual correlation in the detection rate of
lenses as a function of their positions in the image for both the
simulated and the known lens sample. Thus, the completeness of the lens
sample is most likely not significantly affected by whether a lens is
located close to the border or well within the center.

\subsubsection{Volunteer profile}

Here, we investigate if the power\footnote{see the
\Aref{appendix:power} for the meaning of this term and how it is
different from the ``Skill'' defined in \PaperOne} of volunteers is systematically
different between the sample of detected and missed lenses.

We check how the posterior probability $P$ (see \PaperOne for the
mathematical definition) of an image or a subject to contain a lens
changes as the image receives more classifications from multiple
volunteers. A graphical representation of changing probabilities for
increasing classifications is called a trajectory plot. In
\Fref{fig:detmis}, we show the trajectory plots of a few examples of
detected lenses (top left panels) and missed lenses (bottom left panels)
from \StageOne of \sw. The number of classifications (Nclass) for a
subject increase from top to bottom.  Also, every subject is assigned a
prior probability $P_0=2\times10^{-4}$ (grey dashed line) and starts at
the middle of the trajectory plot. The $P$ value of a subject is updated
with every classification from the volunteer.  If a volunteer identifies
a lens candidate, the trajectory moves to the right otherwise moves to
the left. A subject is accepted if it crosses the blue-dashed line
marking the ($P_{\rm acc thresh}=0.95$) on the right and is rejected, if
it crosses the red-dashed line marking the $P_{\rm rej thresh}=10^{-7}$
on the left.

The amount by which the posterior probability $P$ value of a subject will
change depends on how well the volunteers are performing on the training
sample and its current probability. Thus, for a given current
probability, high power volunteers will change the $P$ by a large factor
compared to the low power volunteers. This is evident in the trajectory
plots as large and small distances in between the consequent points
which we refer to as kicks. Comparison of the kick sizes between the
detected and the missed lenses suggests that the missed lenses do not
have as many volunteers giving large kicks. We also note that most of
the large kicks seen in the trajectories of the missed lenses seem to be
moving the subjects to the right. In other words, the high power
volunteers are mostly classifying them as subjects with lens candidates.
Further qualitative comparison and discussion of the trajectory plots is
given in \Aref{appendix:traj}.

For a quantitative comparison of the large and small kicks for the
entire samples of detected and missed known lenses, we show a plot of
histogram on the right of \Fref{fig:detmis}. Qualitatively, there are
four types of volunteers making classifications - those causing positive
large kicks (correct classifications by high power volunteers), those
causing positive small kicks (correct classifications by low power
volunteers), those causing negative small kicks (incorrect
classifications by low power volunteers) and those causing negative
large kicks (incorrect classifications by high power volunteers). The
four histograms in the figure correspond to these four types of
volunteers for each sample (that is, detected or missed). The kick size
is small if $\Delta log{P} = \log{P_{\rm current}} - \log{P_{\rm
previous}} < \Delta\log{P}_{\rm cut}$ (chosen as 1.2) and is large
if greather than $\Delta\log{P}_{\rm cut}$.

Some of the key differences are i) the ratio of positive large kicks to
positive small kicks for the detected sample is higher than the missed
sample suggesting that the fraction of high power volunteers is large
for the detected sample ii) in terms of Nclass, the missed sample is
dominated by volunteers causing negative small kicks whereas for the
detected sample there is a comparable contribution from three types of
volunteers - positive small and large kicks and negative small kicks.
iii) the Nclass from volunteers with negative large kicks are lower, both
for the detected and the missed samples. This is consistent with our
expectation that high power volunteers should not be making incorrect
classifications. Therefore, lack of high power volunteers classifying
the missed lenses seems to be one of the major factors.

We rerun SWAP for \StageOne where we use classifications from volunteers
who produce $\Delta \log{P} > 1.2$ only. This obviously means reducing
the total Nclass per subject by a large fraction. As a result, we also
need to change the $P_{\rm acc thresh}$ which is chosen to be 0.1 and we
find that about a third of the missed lenses are detected while all the
previously detected lenses remain detected too. The others simply do not
have enough classifications from volunteers producing large positive
kicks. Thus, it may be possible to detect more promising candidates if
we preferentially show the rejected systems to a few high power
volunteers. We understand that changing the rejection and acceptance
thresholds will likely increase the rate of FPs along with improved
completeness. After testing on a subsample of both \StageOne and
\StageTwo FPs, we find that a small fraction of FPs are rejected when
only classifications with $\Delta \log{P}>1.2$ only are considered.
Therefore, we think it is worth adopting the new strategy in the future
runs of SWAP in future lens searches.

@@

\begin{figure*}
\begin{center}
\includegraphics[scale=0.6]{det_mis_traj.pdf}
\includegraphics[scale=0.6]{det_mis_hist.pdf}
\caption{ \label{fig:detmis}
Examples of trajectories of some known lenses (top-left: detected and
bottom-left: missed) at \StageOne and histogram of the known lenses for
different types of volunteers (right). Among the sample of missed
lenses, most classifications are from the low power volunteers
identifying these images incorrectly (i.e. $-\Delta$~logP) which
overshadows the small number of correct identifications (i.e.
$\Delta$~logP) coming from both the less and the high power volunteers
combined. For the detected lens sample, most classifications are correct
identifications coming from both low and high power volunteers.
}
\end{center}
\end{figure*}

>>>>

\subsubsection{Blind lens search}

The efficiency of a visual search could potentially
vary in different sections of an
image. Our eyes tend to focus usually at the center of an image and lens
candidates close to the borders could go undetected. Therefore, it is
important to check whether \sw could be missing some of the known lenses
because they happen to be close to the borders of the image cutouts.

@@

We do not find any strong correlation in the detection rate of
lenses as a function of their positions in the image, for either the
simulated or the known lens sample. Thus, the completeness of the lens
sample is most likely not significantly affected by lenses
located close to the image borders.

\subsubsection{Classification Uncertainty}

Each \sw image classification is based on the markers placed by around 10
volunteers (on average, \PaperOne). This number could be
small enough to introduce some scatter in the system performance, arising from
the variations between the small groups of people inspecting each subject.
In \PaperOne we investigated the system performamce in terms of the ``Skill''
of each agent; in \Aref{appendix:power} we define a complementary property,
the ``Power'' of a classification to make a large difference in the probability
$P$ of an image containing a lens. Here, we investigate whether the distribution
of classification power is systematically
different between the detected and missed lenses.

We check how the posterior probability $P$ (see \PaperOne for the
mathematical definition) of an image or a subject to contain a lens
changes as the image receives more classifications from multiple
volunteers. A graphical representation of changing probabilities for
increasing classifications is called a trajectory plot. In
\Fref{fig:detmis}, we show the trajectory plots of a few examples of
detected lenses (top row of panels) and missed lenses (bottom row of panels)
from \StageOne of \sw. Every subject is assigned a
prior probability $P_0=2\times10^{-4}$ (grey dashed line) and starts at
the middle of the trajectory plot. The number of classifications (Nclass) for a
subject increases from top to bottom (subjects move down the trajectory plots
as they are classified).  The $P$ value of a subject is updated
with every classification from the volunteer.  If a volunteer identifies
a lens candidate, the trajectory moves to the right otherwise moves to
the left. A subject is accepted if it crosses the blue-dashed line
marking the ($P_{\rm acc thresh}=0.95$) on the right. It is rejected if
it crosses the red-dashed line marking the $P_{\rm rej thresh}=10^{-7}$
on the left.

The amount by which the posterior probability $P$ value of a subject will
change depends on how well the volunteers are performing on the training
sample and its current probability. Thus, for a given current
probability, some volunteers will change the $P$ by a large factor
compared to others. This is evident in the trajectory
plots, which show both large and small distances between consecutive points
which we refer to as kicks. Comparison of the kick sizes between the
detected and the missed lenses suggests that {\it the missed lenses do not
have as many volunteers giving large kicks}. We also note that most of
the large kicks seen in the trajectories of the missed lenses seem to be
moving the subjects to the right. In other words, certain ``high power''
volunteers are mostly classifying them as subjects with lens candidates.
Further qualitative comparison and discussion of the trajectory plots is
given in \Aref{appendix:traj}.

For a quantitative comparison of the large and small kicks for the
entire samples of detected and missed known lenses, we show a plot of
histogram on the right of \Fref{fig:detmis}. Qualitatively, there are
four types of volunteers making classifications: those causing large, positive
kicks (correct classifications by high power volunteers), those
causing positive small kicks (correct classifications by low power
volunteers), those causing negative small kicks (incorrect
classifications by low power volunteers) and those causing negative
large kicks (incorrect classifications by high power volunteers). The
four histograms in the figure correspond to these four types of
volunteers for each sample (that is, detected or missed). In this plot,
the kick size
is defined as small if
$\Delta log{P} = \log{P_{\rm current}} - \log{P_{\rm
previous}} < \Delta\log{P}_{\rm cut}$ (chosen as 1.2) and is large
if greather than $\Delta\log{P}_{\rm cut}$.

Some of the key differences are as follows.  i) The ratio of  large positive
kicks to small positive kicks for the detected sample is higher than in the
missed sample, suggesting that the fraction of high power volunteers is large
for the detected sample.  ii) The number of classifications received by the
missed sample is dominated small negative kicks. In contrast, for the detected
sample there are comparable contributions from all three of small, positive
kicks, large, positive kicks, and small, negative kicks. iii) The number of
classifications providing large negative kicks is lower for both the detected
and the missed samples. This is consistent with our expectation that high power
volunteers should not be making incorrect classifications.
We conclude that one of the major factors in the \SW system missing lenses is
a lack of high power classifications.

As a demonstration, we re-ran SWAP for \StageOne using only  classifications
that produced large kicks, with $|\Delta \log{P}| > 1.2$.  This obviously meant
reducing the total number of classifications per subject by a large fraction.
As a result, we also needed to change the $P_{\rm acc thresh}$. Choosing this to
be 0.1, we found that about a third of the lenses that had previously been
missed were now detected, while all the previously detected lenses were again
detected. The remaining missed lenses simply do not have enough classifications
from volunteers producing large positive kicks. This experiment shows that it
may be possible to increase the  \sw completeness  by preferentially showing
certain rejected systems -- those that  had never received a high power
classification --  to  volunteers capable of making such classifications.

Changing the rejection and acceptance
thresholds will likely increase the rate of false positives along with improved
completeness.
% After testing on a subsample of both \StageOne and
% \StageTwo FPs, we find that a small fraction of FPs are rejected when
% only only classifications with $|\Delta \log{P}|>1.2$ are considered.
% Therefore, we think it is worth adopting the new strategy in the future
% runs of SWAP in future lens searches.
This will need to be further quantified before detailed recommendations can be
made. However, dynamically assigning certain subjects to volunteers according to
various measures of their skill seems like a fruitful line of investigation when
seeking to improve the system performance in future.

@@

\begin{figure*}
\begin{center}
\includegraphics[scale=0.6]{det_mis_traj.pdf}
\includegraphics[scale=0.6]{det_mis_hist.pdf}
\caption{ \label{fig:detmis}
Left: Examples of \StageOne trajectories of some known lenses.
Upper panels show detected
lenses, lower panels show missed lenses (false negatives). Right:
Histogram of classification power received by known lenses.
Among the sample of missed
lenses, most classifications are from the low power volunteers
identifying these images incorrectly (i.e. $-\Delta$~logP) which
overshadows the small number of correct identifications (i.e.
$\Delta$~logP) coming from both the less and the high power volunteers
combined. For the detected lens sample, most classifications are correct
identifications coming from both low and high power volunteers.
}
\end{center}
\end{figure*}
